\chapter{Referencias}

\section*{Estándares y manuales}
\begin{itemize}
    \item NASA-SP-2007-6105 Rev2 — \textit{NASA Systems Engineering Handbook}, 2007.
    \item ECSS-E-ST-10C Rev1 — \textit{Space Engineering: System Engineering General Requirements}, ESA, 2017.
    \item ECSS-E-ST-10-02C Rev1 — \textit{Verification}, ESA, 2018.
    \item ECSS-E-ST-10-03C Rev1 — \textit{Testing}, ESA, 2022.
    \item ECSS-E-ST-20C Rev2 — \textit{Electrical and electronic}, ESA, 2022.
    \item ECSS-E-ST-31C — \textit{Thermal control general requirements}, ESA, 2008.
    \item ECSS-M-ST-40C Rev1 — \textit{Project Management}, ESA, 2008.
    \item GSFC-STD-7000A — \textit{General Environmental Verification Standard (GEVS)}, NASA Goddard, 2019.
    \item CubeSat Design Specification Rev~14 — Cal Poly SLO, 2020.
    \item CCSDS 133.0-B-2 — \textit{Space Packet Protocol}, CCSDS, 2020.
    \item ISO 21502:2020 — \textit{Project, programme and portfolio management}, ISO.
    \item PMBOK Guide, 7\textsuperscript{th} Edition — PMI, 2021.
\end{itemize}

\section*{Documentos de referencia técnica}
\begin{itemize}
    \item Wertz, J. R., Everett, D. F., \& Puschell, J. J. — \textit{Space Mission Engineering: The New SMAD}, Microcosm Press, 2011.
    \item Larson, W. J. \& Wertz, J. R. — \textit{Space Mission Analysis and Design}, 3\textsuperscript{rd} ed., 1999.
    \item SpaceLab UFSC — \textit{GOLDS-UFSC (FloripaSat-2) Design and Mission Overview}, junio 2023. Disponible en \texttt{https://github.com/spacelab-ufsc/floripasat2-doc}.
\end{itemize}

\section*{Documentos internos INTISAT (este repositorio)}
\begin{itemize}
    \item \texttt{03\_recursos/APSCO/APSCO\_Invitacion\_Summer\_Camp\_Xian\_2026.pdf}
    \item \texttt{03\_recursos/APSCO/GUIA\_estandares\_y\_referencias.md}
    \item \texttt{03\_recursos/presupuestos/Requerimiento de energía - INTISAT.xlsx}
    \item \texttt{01\_CDRs\_subsistemas/} — CDRs detallados por subsistema (OBC, UHF, EPS, Microscopía).
\end{itemize}
